---
title: "Coding Efficiency Demo"
author: "Jimmy Zhang"
date: "2022-10-17"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. What Is Coding Efficiency?

Efficiency means "working well to avoid wasting energy, efforts, time, money, etc." It has a formal definition: $$\eta = \frac{W(Work)}{Q(Effort)}.$$ In words, it means how much work has been done in a unit of effort. A high efficiency means more work has been done in a unit of effort. 

Two types of efficiency are defined:
    
1. *Algorithmic Efficiency*: how quickly the computer can undertake a piece of work given a particular piece of code.
    
2. *Programmer Productivity*: the amount of useful work a person **(not a computer)** can do per unit time.

We must make a balance between algorithmic efficiency and programmer productivity: It is possible that we may pay too much human efforts to optimize an algorithm (e.g., spend 50 hours to make a function generalizable). The works that we complete for this optimization may seem not enough given the efforts we make, although we complete a more efficient algorithmic. 

# 2. Why can we achieve efficient programming in R? 

R has some significant advantages:

* R is not compiled but it calls compiled code. It removes the laborious stage of compiling your code before being able to run it.

* R is a functional and object orientated language. It is possible to write complex and flexible functions in R that get a huge amount of work done with a single line of code.

```{r eval=FALSE, include=FALSE}
compute_median <- function (vec) {
  sorted_vec <- sort(vec, decreasing = TRUE)
  if (length(vec) %% 2 == 0) {
    output <- (sorted_vec[length(sorted_vec)/2] + sorted_vec[length(sorted_vec)/2 + 1])/2
  }
  else {
    output <- sorted_vec[ceiling(length(sorted_vec)/2)]
  }
  print(output)
}

compute_median(c(4,5,99,45))
```

Object-oriented programming (OOP): a programming paradigm that relies on the concept of classes and objects. It is used to structure a software program into simple, reusable pieces of code blueprints (usually called classes), which are used to create individual instances of objects. Python, Java, C++ are all examples of OOP languages. 

![An illustration of an OOP example.]("/Users/jimmy_z/R Projects/CodingEfficiency/images/OOP_example"){width=50%}

* R uses RAM for memory.

RAM (random access memory) is a computer's short-term memory, where the data that the processor is currently using is stored. Your computer can access RAM memory much faster than data on a hard disk, SSD, or other long-term storage device, which is why RAM capacity is critical for system performance. 

![An illustration of RAM.]("/Users/jimmy_z/R Projects/CodingEfficiency/images/RAM"){width=50%}

* R is supported by excellent Integrated Development Environments (IDEs).

An integrated development environment is a software application that provides comprehensive facilities to computer programmers for software development. An IDE normally consists of at least a source code editor, build automation tools and a debugger. 

![An illustration of IDE.]("/Users/jimmy_z/R Projects/CodingEfficiency/images/IDE"){width=50%}

* R has a strong user community.
If you encounter a problem that has not yet been solved, you can simply ask the community:
  - StackOverflow
  - R-bloggers
  
# 3. Some basic skills and conventions to promote efficient coding in R

1. Conventions

* Packages names are in bold: **lavaan**.
* Functions and R objects are in a code font: `ggplot()`, `x`

2. Benchmark

* It allows you to keep track of the performce of your function: **microbenchmark**

```{r}
library("microbenchmark")
df = data.frame(v = 1:4, name = letters[1:4])
microbenchmark(df[3, 2], df[3, "name"], df$name[3])
```

3. Profiling

* As R users, we may want our codes to run faster. However, it’s not always clear how to accomplish this. 

* Profiling is about testing large chunks of code. It is able to identify bottlenecks in your R scripts.

```{r}
library("profvis")
profvis(expr = {
  
  # Stage 1: load packages
  library("rnoaa") # not necessary as data pre-saved
  library("ggplot2")
  
  # Stage 2: load and process data
  out = readRDS("extdata/out-ice.Rds")
  df = dplyr::rbind_all(out, id = "Year")
  
  # Stage 3: visualise output
  ggplot(df, aes(long, lat, group = paste(group, Year))) +
    geom_path(aes(colour = Year)) 
  ggsave("figures/icesheet-test.png")
}, interval = 0.01, prof_output = "ice-prof")
```

# 4. Efficient Programming in R

## 1. Some broad but fundamental ideas

Ultimately calling an R function always ends up calling some underlying *C/Fortran code*, which are low level languages that demand more from the programmer. They force you to declare the type of every variable used, give you the burdensome responsibility of memory management and have to be compiled.

For example, if you call a base R function `runif`, you can view the source code of this function: 

![An illustration of runif.]("/Users/jimmy_z/R Projects/CodingEfficiency/images/runif"){width=50%}

We can see that `runif` only contains a single line that calls `C_runif()`. With this being said, R is a higher level language that somehow "simplifies" our coding process. 

A golden rule in R programming is to access the underlying C/Fortran routines as quickly as possible: the fewer functions calls required to achieve this, the better.

```{r}
x <- c(1, 2, 3, 4, 5)

# Method 1:

x <- x + 1

# Method 2:

for(i in seq_len(length(x))) 
  x[i] = x[i] + 1 

microbenchmark(x <- x + 1, 
for(i in seq_len(length(x))) 
  x[i] = x[i] + 1 )
```

Another useful rule is to pre-allocate your vector and then fill in values. A best way to avoid massive computation is to avoid building an object (e.g., vector, data frame, list, etc).

```{r}
method1 = function(n) {
  vec = NULL # Or vec = c()
  for(i in seq_len(n))
    vec = c(vec, i)
  vec
}

method2 = function(n) {
  vec = numeric(n)
  for(i in seq_len(n))
    vec[i] = i
  vec
}

method3 = function(n) seq_len(n)
```

Method 1 creates a null vector that contains nothing. 

```{r}
vec = NULL
length(vec)
```

Then it adds values in this vector and builds up the vector gradually. 

Method 2, however, creates a an object of the final length and then changes the values in the object by subscripting.

```{r}
n = 100
vec = numeric(n)
length(vec)
```

Method 3 directly creates the final object.

It is not difficult to guess which one wins in terms of efficiency.

![An illustration of Three Methods Comparison.]("/Users/jimmy_z/R Projects/CodingEfficiency/images/method_comp"){width=50%}

## 2. Vectorization

Given the faster speed of accessing the underlying C/Fortran routines, we can reduce the number of function calls required by working on vectorization. 

```{r}
# Using for loop
log_sum = 0
for(i in 1:length(x))
  log_sum = log_sum + log(x[i])

# Using vectorized method
log_sum = sum(log(x))
```

One of my own challenges is that I particularly like `for loop` and `while loop` to construct complex functions because this is one of the simplest ways to get it started. Although this way of thinking works, sometimes it'll greatly increase the consumption of computation works. We can spend a little bit more time to challenge ourselves: how to make it simpler by vectorization? 











  



 