---
title: "Exercises"
author: "Jimmy Zhang"
date: "2022-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Vectorization

1. Suppose we have a large-scale survey dataset with 50,000 respondents, who self-reported whether they voted in a recent election and their age.

```{r}
n <- 100000
set.seed(123)
d <- data.frame(
  id = 1:n,
  turnout = sample(c("voted", "did not vote"), 
                   size = n, replace=TRUE),
  age = sample(18:100, size=n, replace=TRUE)
)
```

## 1. Count how many respondents voted:

Here I have an example code to achieve this task using `for loop`:

```{r}

func1 <- function(){
  
  count <- 0
  
  for (i in 1:nrow(d)){
    if (d$turnout[i]=="voted"){
      count <- count + 1
    }
  }
  return(count)
}

system.time(count <- func1())

```

How can we improve this function and make it faster?

```{r}
func2 <- function(){

}

system.time(count <- func2())
```

## 2. Compute the average age of respondents who voted:

```{r}
func1 <- function(){
  voters_age <- c()
  for (i in 1:nrow(d)){
    if (d$turnout[i]=="voted"){
      voters_age <- c(voters_age, d$age[i])
    }
  }
  return(mean(voters_age))
}
system.time(count <- func1())
```

Your solution:

```{r}
func2 <- function(){

}
system.time(func2())
```

## 3. Take a random sample of 200 rows in the dataset that correspond to respondents who did not vote

```{r}
func1 <- function(){
  
  # sample random uniform number
  d$random <- runif(nrow(d))
  # sort by this random number
  dd <- d[order(d$random),]
  # loop over rows using 'while'
  smp <- data.frame()
  i <- 1
  while (nrow(smp)<200){
    if (d$turnout[i]=="did not vote"){
      smp <- rbind(smp, d[i,])
    }
    i <- i + 1
  }
  return(smp)

}

system.time(smp <- func1())

```

Your solution here:

```{r}
func2 <- function(){

}
system.time(func2())
```

# Algorithm Complexity

Write a function that returns all prime numbers from 2 until an arbitrary number *n*. 

A couple of clues to get you started: (1) remember that a prime number can only be divided by 1 and itself; (2) in R, you can compute the modulo (aka remainder) of a number divided by another as `10 %% 5` -- this will return 0 because when you divide 10 by 5 there is no remainder. In contrast, `11 %% 5` will return 1.

Can you write a solution that takes less than 1 second to identify all prime numbers for n=10,000?





